Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-31 algorithm.tex
AA: during the answer revision

With regard to the performance of the proposed approach we can estimate that the
performance of $PS$ is strictly related to the total number of
different configurations evaluated during the exploration. We can
safely make such assumption, because the process of evaluating a new
configuration is the time-dominating one among all the operations
performed. As we will show in the
next Section, this is especially true in the hardware/software
design space exploration scenario we are addressing, since benchmark
compilation and execution times
are orders of magnitude bigger than the splitting/merging operations and
innovation scores computations.  Ideally, the $i$-th era would provide a new set
$test_i$ of $K$ configurations to be tested. However, we cannot know
a priori how many of
these will actually be new configurations, i.e. requiring a process of
evaluation. Thus, number of total new simulations $N_{tot}$ 
can be estimated to be bounded as follows:
\[
K \le N_{tot} \le K \times E_{tot}
\]
where $E_{tot}$ is the number of eras executed and assuming that at
least $K$ evaluation should be performed in order to obtain the first
pareto $\mathscr{P}_0$.

>>>>

With regard to the performance of the proposed approach, we observe that the
performance of $PS$ is related to the total number of
different configurations evaluated during the exploration. We can
safely make such assumption, because the process of evaluating a new
configuration is the time-dominating one among all the operations
performed. As we will show in the
next Section, this is especially true in the hardware/software
design space exploration scenario we are addressing, since benchmark
compilation and execution times
are orders of magnitude bigger than the splitting/merging operations and
innovation scores computations.  The $i$-th era would provide a new set
$test_i$ of $K$ configurations to be tested. However, we cannot know
a priori how many of
these will actually be new configurations, i.e. requiring a process of
evaluation. Thus, the number of total new simulations $N_{tot}$ 
is bounded as follows:
\[
K \le N_{tot} \le K \times E_{tot}
\]
where $E_{tot}$ is the number of eras executed and assuming that at
least $K$ evaluation should be performed in order to obtain the first
pareto $\mathscr{P}_0$.

### 2014-10-31 appendix.tex
AA: during the answer revision

\subsection{Other metrics}
Un'altra interessante metrica potrebbe essere questa: Auhtors of \cite{zitzler_ec00} remark that a desirable property of the Pareto front is the spread of the non-dominated configurations all over the configuration space. In other words, the more similar the non-dominated configuration distribution is to the uniform distribution, the better is the quality of the Pareto front, since we have not been captured by only one possible subset of Pareto configurations but we are able to span all over the configuration space and find configurations whose parameter values differ considerably. Mi aspetto che PS performi bene sotto questo punto di vista. Ovviamente, non so se e' il caso di considerarla ora. Se non e' il caso, puoi spostare questo testo nel file ``appendix.tex'' come possibile spunto futuro. Per quantificare questa cosa, si potrebbe usare come metrica la normalised absolute dispersion error pero' non nello spazio degli obiettivi, ma nello spazio delle configurazioni.

>>>>

\subsection{Other metrics}
Per rappresentare l'extent del Pereto front, forse
sarebbe meglio evitare di usare il variation range, per come l'hai
definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5
``Metrics of performance'', che mi pare rappresenti la stessa cosa,
concettualmente. Questo ci permette di essere meno attaccabili (del
tipo ``non te la prendere con noi, prenditela con Zitzler (che e'
citato 3064 volte!) se non capisci l'utilita' della metrica'').

Un'altra interessante metrica potrebbe essere questa: Auhtors of \cite{zitzler_ec00} remark that a desirable property of the Pareto front is the spread of the non-dominated configurations all over the configuration space. In other words, the more similar the non-dominated configuration distribution is to the uniform distribution, the better is the quality of the Pareto front, since we have not been captured by only one possible subset of Pareto configurations but we are able to span all over the configuration space and find configurations whose parameter values differ considerably. Mi aspetto che PS performi bene sotto questo punto di vista. Ovviamente, non so se e' il caso di considerarla ora. Se non e' il caso, puoi spostare questo testo nel file ``appendix.tex'' come possibile spunto futuro. Per quantificare questa cosa, si potrebbe usare come metrica la normalised absolute dispersion error pero' non nello spazio degli obiettivi, ma nello spazio delle configurazioni.

### 2014-10-31 experiments.tex
AA: during the answer revision

Although an analogy between the concepts of ``era'' and ``generation''
can be considered, comparisons have been carried out using as
reference a fixed amount of total unique simulations, instead of a given
number of eras/generations . The purpose is to make the comparison
fair and accurate from the algorithm performance perspective, since a fixed budget of
different simulations is more directly related to the actual amount of time
required to execute the exploration; on the other side, simply
referring to a given number of generations/eras could be inaccurate
when a significative number of configuration remain unaltered in the
current population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted, similar to the one shown in
Figure~\ref{fig:kimpact} which refer to the alloca\_test application
and a fixed amount of 500 configurations. We found that initially increasing values
of $K$ seem to improve both Pareto (as in the picture on the left) and objective
distribution (on the right), but after some threshold value the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value of $K$ considered (100). Similar trends have been
found for the other benchmarks (not included for space reasons),
showing on average a value of $K$ around 30 as
reasonable sweet spot for the objectives. A complete resource with the
whole set of results and data obtained in the tests performed can be found online
at~\cite{ps_results}.

\begin{table}
	\tbl{Algorithm Parameters\label{tab:defaults}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Default Value} \\
	\hline
	GA population size & 30 \\
	GA crossover probability & 0.8 \\
	GA mutation probability  & 0.1 \\
	$K$ era budget & 30 \\
	alpha threshold & 1 \\
	Total simulations budgets & 100-500 \\
	\hline
	\end{tabular}}
\end{table}

\begin{figure}
  \figLC{kimpact}{Impact of increased K values on Pareto fronts and distribution for the alloca\_test application.}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K.eps} 
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K_ALL.eps} 
  \end{center}
\end{figure}

@@

\comment{AA: Per rappresentare l'extent del Pereto front, forse
sarebbe meglio evitare di usare il variation range, per come l'hai
definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5
``Metrics of performance'', che mi pare rappresenti la stessa cosa,
concettualmente. Questo ci permette di essere meno attaccabili (del
tipo ``non te la prendere con noi, prenditela con Zitzler (che e'
citato 3064 volte!) se non capisci l'utilita' della metrica'').}

@@

\figR{dispersion}(c) shows the average normalized absolute dispersion
errors for different benchmarks for PS and GA exploration.  As it can
be observed, GA shows a tendency in discovering more uniformly
dispersed points (i.e. higher values in the chart). These results,
together with the variation ranges of \figR{dispersion}(a) and (b)
confirm the of density vs extension trade-offs when comparing PS and
GA approaches. Again, it is worth pointing out how a definitive claim
about which algorithm shows better results, being that strictly
dependent on the design requirements and constraints, e.g.  whether we
need wide or dense results.
\comment{Da lasciare, da spostare ? da mostrare in modo semplice ?  non so}

>>>>

Although an analogy between the concepts of ``era'' and ``generation''
can be considered, comparisons have been carried out using as
reference a fixed amount of total unique simulations, instead of a given
number of eras/generations . The purpose is to make the comparison
fair and accurate from the algorithm performance perspective, since a fixed budget of
different simulations is more directly related to the actual amount of time
required to execute the exploration; on the other side, simply
referring to a given number of generations/eras could be inaccurate
when a considerable number of configurations remain unaltered in the
current population (e.g. because of a local optimum)\comment{AA: Da 'on the other side' fin qui non ho capito cosa vuoi dire}.

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, a sensitivity analysis on the algorithm parameters is beyond the scope of this work. Nevertheless, we conducted several tests with
different values of $K$, similar to the ones shown in
Figure~\ref{fig:kimpact} which refers to the alloca\_test application
and a total budget of 500 configurations. We found that initially increasing values
of $K$ seems to improve both Pareto (as in the picture on the left) and objective
distribution (on the right), but after some threshold value the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value of $K$ considered (100). Similar trends have been
found for the other benchmarks (not included for space reasons),
showing on average a value of $K$ around 30 as
reasonable sweet spot for the objectives. A complete resource with the
whole set of results and data obtained in the tests performed can be found online
at~\cite{ps_results}.

\begin{table}
	\tbl{Algorithm Parameters\label{tab:defaults}}{
	\begin{tabular}{cll}
	\hline
	\multicolumn{1}{c}{Algorithm} & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Default Value} \\
	\hline
	GA & population size & 30 \\
	GA & crossover probability & 0.8 \\
	GA & mutation probability  & 0.1 \\
	PS & era budget ($K$) & 30 \\
	PS & innovation threshold ($\alpha$ ) & 1 \\
	both PS and GA & Total simulations budget & 100-500 \\
	\hline
	\end{tabular}}
\end{table}

\begin{figure}
  \figLC{kimpact}{Impact of increased K values on Pareto fronts and distribution for the alloca\_test application. \comment{AA: non capisco cosa sia la figura a destra. Inoltre, io vedo sulle ascisse e le ordinate di entrambi i grafici, le stesse grandezze (Time vs Power), quindi mi aspetto di trovare esattamente lo stesso plot sia a destra che a sinistra}}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K.eps} 
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K_ALL.eps} 
  \end{center}
\end{figure}

@@

@@

\figR{dispersion}(c) shows the average normalized absolute dispersion
errors for different benchmarks for PS and GA exploration.  As it can
be observed, GA shows a tendency in discovering more uniformly
dispersed points (i.e. higher values in the chart). These results,
together with the variation ranges of \figR{dispersion}(a) and (b)
confirm the of density vs extension trade-offs when comparing PS and
GA approaches. We highlight that PS shows a stronger \emph{uniform convergence} (discussed in \secL{metrics}), i.e. it provides a more extended Pareto front, avoiding the designer to neglect some important options, and offering a wider view of the potentiality of the designed device.

### 2014-10-31 introduction.tex
AA: during the answer revision

Authors of \cite{dellnitz2005covering,dellnitz2009computation} propose
three different methods for the numerical approximation of the Pareto
set. The first two require the knowledge of the gradient of the
objective functions, which is usually not available in embedded system
design, since the closed form expression of the objective functions is
not known a priori. The third one, called ``Sampling algorithm'', does
not need this requirement and is the closest to the PS approach
presented in this work, although the
mechanism is different. At every iteration a set of regions of
configuration space is constructed with the aim to cover the Pareto
set. The size of these regions becomes smaller at each iteration so
that the covering becomes tighter. A random set of test points are
evaluated in each new region; only the regions containing
non-dominated points will be considered in the next iteration.
However, an important feature differentiates it from PS. The Sampling algorithm only the
selected regions are examined, even though there is no guarantee that
all the Pareto points are included in them. As a consequence, the
Sampling algorithm may erroneously and prematurely neglect some
regions that may possibly contain a non-negligible number of Pareto
points. The PS algorithm avoids this, since even though a region has not
provided non-dominated points, it is not neglected, rather it is
simply populated with fewer test points, thus giving the chance to
find non-dominated points in them. From this behavior also comes
another important difference. The Sampling algorithm of
\cite{dellnitz2005covering} keeps reinforcing the analysis of the
regions selected in the previous iterations. In other words, it keeps
zooming into the regions that have been considered ``interesting'',
i.e. rich of non-dominated points, in the previous iterations. The
philosophy behind our algorithm is different. A region, which has been
considered interesting at a certain iterations, may be not in the
future, since it has been thoroughly analyzed. On the other hand, a
region, initially uninteresting, may turn out to be interesting in the
future, once all the other regions have been already exploited.
Therefore, from an iteration to the next, our algorithm zoom in on
regions that are currently interesting and zoom out from the
uninteresting ones,  spanning in a broader and more dynamical way the
configuration space. In addition, in the Sampling region only the
number of non-dominated points is accounted to see how interesting a
region is. On the contrary, we also consider the \emph{innovation}, as
already anticipated in the introduction.

>>>>

Authors of \cite{dellnitz2005covering,dellnitz2009computation} propose
three different methods for the numerical approximation of the Pareto
set. The first two require the knowledge of the gradient of the
objective functions, which is usually not available in embedded system
design, since the closed form expression of the objective functions is
not known a priori. The third one, called ``Sampling algorithm'', does
not need this requirement and is the closest to the PS approach
presented in this work, although the
mechanism is different. At every iteration a set of regions of
configuration space is constructed with the aim to cover the Pareto
set. The size of these regions becomes smaller at each iteration so
that the covering becomes tighter. A random set of test points are
evaluated in each new region; only the regions containing
non-dominated points will be considered in the next iteration.
However, an important feature differentiates it from PS. In the Sampling algorithm only the
selected regions are examined, even though there is no guarantee that
they contain all the Pareto points. As a consequence, the
Sampling algorithm may erroneously and prematurely neglect some
regions that may possibly contain a non-negligible number of Pareto
points. The PS algorithm avoids this, since even though a region has not
provided non-dominated points, it is not neglected, but it is
populated with fewer test points, allowing to
find its other non-dominated points. From this behavior also comes
another important difference. The Sampling algorithm of
\cite{dellnitz2005covering} keeps reinforcing the analysis of the
regions selected in the previous iterations. In other words, it keeps
zooming in into the regions that have been considered ``interesting'',
i.e. rich of non-dominated points, in the previous iterations. The
philosophy behind our algorithm is different. A region, which has been
considered interesting at a certain iterations, may be not in the
future, since it has been thoroughly analyzed. On the other hand, a
region, initially uninteresting, may turn out to be interesting in the
future, once all the other regions have been already exploited.
Therefore, from an iteration to the next, our algorithm zoom in on
regions that are currently interesting and zoom out from the
uninteresting ones,  spanning in a broader and more dynamical way the
configuration space. In addition, in the Sampling region only the
number of non-dominated points is accounted to see how interesting a
region is. On the contrary, we also consider the \emph{innovation}, as
already anticipated in the introduction.
